---
title: 'The Mathematics behind bedrocktrader: Villager Trade Probabilities in Minecraft: Bedrock Edition'
title-meta: 'The Mathematics behind bedrocktrader: Villager Trade Probabilities in Minecraft: Bedrock Edition'
author:
  - name: Rento Saijo
    corresponding: true
    email: rentosaijo0527@gmail.com
    affiliations:
      - Connecticut College
keywords:
  - 'Minecraft: Bedrock Edition'
  - Villager trading
  - Enchanting
  - Probability
  - R
abstract: |
  Villager trade tables describe random mechanisms rather than a single realized inventory. Turning those tables into probabilities requires more than flattening JSON: source trades are sampled within groups, item choices branch into alternatives, item functions generate additional outcomes, and identical visible offers can arise through several source paths. This note develops the probability model implemented in bedrocktrader for the vanilla Minecraft: Bedrock Edition 1.26.30.5 trade tables. It formalizes context conditioning and the package's trade-to-option-to-offer hierarchy, derives selection probabilities when repeated source entries create multiplicity, and shows how exact offers aggregate into item specifications and trade summaries. Separate sections cover librarian enchanted books, capped treasure prices, fully unlocked librarian probabilities, and the documented model used for enchanted equipment. Worked examples recover a 2.79% probability of a fully unlocked librarian offering Mending for at most 26 emeralds and exact Efficiency II diamond-axe probabilities of 6.07% for a toolsmith and 12.13% for a weaponsmith. The model is evaluated analytically and bundled with the package; no simulation, resampling, network access, or runtime JSON parsing is required.
date: 2026-07-26
bibliography: references.bib
number-sections: true
---

## Introduction

The vanilla Minecraft: Bedrock Edition trade files look deceptively close to
ordinary tabular data. A profession has five tiers, each tier contains one or
more groups, and each group contains possible trades. Yet a row in the source
does not necessarily describe a row on a player's trading screen. Groups select
subsets of trades, `choice` arrays replace one item with one of several
alternatives, contextual filters depend on the villager's biome variant or
dimension, and functions may determine an auxiliary value, map destination,
potion, enchantment, or price. Three visibly identical red-candle source
entries, for example, represent three chances to obtain the same displayed
offer.

bedrocktrader converts those mechanisms into flat base R data frames and
associated analysis functions. The package is pinned to Mojang's immutable
Minecraft Bedrock Samples release `v1.26.30.5` [@mojang2026samples]. During
development, the updater retrieves the vanilla trade tables, villager entity
definition, and enchantment registry; verifies each Git blob hash; validates
the supported structures; and writes normalized private objects to
`R/sysdata.rda`. Installed functions read those bundled objects, so runtime
results are offline and reproducible.

This note documents the implemented mathematics. Microsoft specifies the
trade-table hierarchy and meanings of `num_to_select`, `wants`, `gives`,
`max_uses`, and related fields [@microsoft2026trade;
@microsoft2026tradegroup]. Microsoft also describes
`enchant_book_for_trading` and `enchant_with_levels`, although the latter
delegates important details to the game's enchanting algorithm
[@microsoft2025enchanting]. For complete equipment outcomes, bedrocktrader
therefore supplements the pinned Mojang files with the enchanting mechanics
recorded in a pinned Minecraft Wiki revision [@minecraftwiki2026enchanting].
The resulting probabilities are exact under that documented model. They should
not be read as a claim that every undocumented Bedrock engine constant has been
independently verified. Canonical enchantment identifiers and supported level
ranges follow Mojang's pinned registry together with Microsoft's enchantment
documentation [@microsoft2026enchantments].

The central distinction is resolution. A *trade row* summarizes a flattened
source trade and any explicit item-choice combination. An *option row* resolves
the concrete item specification while marginalizing over its possible prices.
An *offer row* resolves both specification and price. These resolutions answer
different questions, and the probability attached to a row always matches what
that row displays.

## Notation and context conditioning

Let $c=(v,d)$ denote the requested villager context, where $v$ is a
villager biome variant and $d$ is a dimension. Most profession tables do not
use either value. The pinned fisherman table contains variant-dependent boat
offers, while the cartographer table uses both variant and dimension filters.
bedrocktrader requires the relevant context instead of assigning Plains or the
Overworld silently.

For group $g$, let $\mathcal{S}_g$ be the multiset of authored source
entries. A filter indicator

$$I_s(c)=
\begin{cases}
1, & \text{when source entry }s\text{ applies under }c,\\
0, & \text{otherwise}
\end{cases}
$$
is evaluated before group sizes or probabilities. The applicable multiset is

$$\mathcal{S}_g(c)=\{s\in\mathcal{S}_g:I_s(c)=1\},
\qquad
n_g(c)=|\mathcal{S}_g(c)|.
$$
This ordering matters. If a group contains several variant-specific
alternatives, counting all entries before filtering would dilute the
probability of the one alternative that can actually occur for the supplied
villager. Package output reports `num_trades = n_g(c)`, counting authored
entries rather than the number of visually distinct rows.

Every applicable source path is represented privately even when two paths
produce the same public trade. Public identifiers then preserve a nested
relationship:

$$\texttt{trade\_id}\longrightarrow
\texttt{option\_id}\longrightarrow
\texttt{offer\_id}.
$$
A `trade_id` identifies one authored trade after identical source entries have
been collapsed with their multiplicity retained. Mojang `choice` arrays can
create several trade-view rows sharing that identifier. An `option_id`
identifies one concrete result specification, such as a complete enchantment
set. Its offers share the same specification but may charge different prices.
Finally, `offer_id` identifies one exact combination of wants, gives,
specification, and price.

## Source-trade selection

### Selection without multiplicity

Suppose group $g$ has $n=n_g(c)$ applicable authored entries and selects
$k$ entries uniformly without replacement. For an ordinary source entry,

$$\Pr(\text{selected}\mid c)=\frac{k}{n},
\qquad 0\leq k\leq n.
$$
Mojang's `num_to_select` supplies $k$. The value `-1` is bedrocktrader's
normalized representation of the pinned select-all convention, in which every
applicable entry has selection probability one. The pinned fisherman master
group is the one known omission: its pufferfish trade and applicable
variant-specific boat trade are both selected, so the updater records
`num_to_select = -1` and rejects additional unexpected omissions.

### Selection with multiplicity

Several authored entries can yield one public trade. Let $w$ be the number
of entries producing that same result. The public row appears if the selected
subset contains at least one of those $w$ entries. Of the
$\binom{n}{k}$ equally likely subsets, $\binom{n-w}{k}$ avoid all $w$
copies. Consequently,

$$p_{\mathrm{trade}}(n,k,w)
=1-\frac{\binom{n-w}{k}}{\binom{n}{k}},
\tag{1}
$$
where the numerator is zero whenever $n-w<k$. Equation (1) reduces to
$k/n$ when $w=1$.

The librarian master candle group provides a compact example. Mojang authored
three identical red-candle entries and one yellow-candle entry, with one of the
four selected. Red therefore has

$$1-\frac{\binom{4-3}{1}}{\binom{4}{1}}
=1-\frac{1}{4}
=\frac{3}{4},
$$
whereas yellow has $1/4$. bedrocktrader returns one red row and one yellow
row because the three red rows would be publicly indistinguishable. It still
reports `num_trades = 4`, and the retained private multiplicity supplies
$w=3$.

## Choices, functions, and row probabilities

A source trade may contain an explicit `choice` array in a want or give
position. If position $r$ contains $c_r$ alternatives and the positions are
resolved uniformly, one concrete cross-product has conditional probability

$$p_{\mathrm{choice}}
=\prod_r\frac{1}{c_r}
=\frac{1}{\prod_r c_r}.
\tag{2}
$$
The source tables can also apply a function to the given item. Let $f$ index
a function outcome and $p_{\mathrm{function}}(f)$ be its conditional
probability. Uniform auxiliary-value functions assign equal probability to
each inclusive value. Fixed potion and exploration-map functions have one
resolved outcome with probability one. Enchantment functions use the models
developed below. `random_dye` remains unresolved because the pinned static
sources do not establish a complete leather-color distribution.

For an exact offer $o$, the implemented factorization is

$$p_{\mathrm{offer}}(o\mid c)
=p_{\mathrm{trade}}(o\mid c)
\times p_{\mathrm{choice}}(o\mid c)
\times p_{\mathrm{function}}(o\mid c).
\tag{3}
$$
The factors in (3) describe sequential parts of the same mechanism. Context
filtering precedes them, source selection chooses a trade, explicit item
choices choose its flat item combination, and the function chooses the
generated result.

### Aggregation across resolutions

Let $\mathcal{F}(a)$ be the exact offers associated with option $a$.
The option probability marginalizes over prices:

$$p_{\mathrm{option}}(a\mid c)
=\sum_{o\in\mathcal{F}(a)}p_{\mathrm{offer}}(o\mid c).
\tag{4}
$$
Let $\mathcal{A}(t,b)$ be the options associated with trade identifier $t$
and one explicit base item-choice combination $b$. The probability shown on
that trade-view row is

$$p_{\mathrm{row}}(t,b\mid c)
=\sum_{a\in\mathcal{A}(t,b)}
 p_{\mathrm{option}}(a\mid c).
\tag{5}
$$
Thus, summing price-specific offers recovers the option row, and summing the
options for a base item choice recovers its trade row. Rows across an entire
tier need not sum to one: separate groups may each contribute an offer, and a
villager can expose several trades simultaneously.

## Librarian enchanted books

### Enchantment and level

The pinned librarian table contains four independent opportunities for an
enchanted book: novice, apprentice, and journeyman book candidates are selected
with probability $1/2$, and the expert candidate is selected with probability
$1/3$. Conditional on selecting the book trade, the documented model chooses
one of $E=39$ villager-attainable enchantments uniformly
[@minecraftwiki2026trading]. Soul Speed, Swift Sneak, and Wind Burst are
recognized by the package registry but excluded from this librarian pool.

Let $e$ be an attainable enchantment and $m_e$ its maximum level. Conditional
on $e$, the level $L$ is uniform on $\{1,\ldots,m_e\}$. Therefore,

$$\Pr(E=e,L=\ell\mid\text{book selected})
=\frac{1}{39}\frac{1}{m_e},
\qquad \ell=1,\ldots,m_e.
\tag{6}
$$
Because different enchantments have different maximum levels, enchantment-level
pairs are not collectively uniform. Aqua Affinity I has conditional
probability $1/39$, while each Efficiency level has conditional probability
$1/(39\cdot5)$.

### Emerald-price generation

Microsoft gives the pinned `enchant_book_for_trading` parameters as
`base_cost = 2`, `base_random_cost = 5`, `per_level_cost = 3`, and
`per_level_random_cost = 10` [@microsoft2025enchanting]. For level $L$, define

$$R_L=5+10L,
\qquad
U_L\sim\operatorname{DiscreteUniform}\{0,\ldots,R_L-1\}.
$$
The ordinary price before treasure adjustment is

$$C_0(L)=2+3L+U_L.
\tag{7}
$$
Hence the normal supports are $5$--$19$, $8$--$32$, $11$--$45$,
$14$--$58$, and $17$--$71$ before the universal stack cap. The package
applies the 64-emerald cap to final prices, so level V ordinary prices above 64
also merge at 64. For a treasure enchantment,

$$C_T(L)=\min\{2C_0(L),64\}.
\tag{8}
$$
The support below 64 is therefore even and can be noncontiguous inside the
minimum--maximum interval shown at option resolution.

Let $h_T(x)=\min(2^T x,64)$, where $T$ is the treasure indicator. Several
underlying values can map to the same capped price. For displayed emerald
price $c$,

$$\Pr(C=c\mid e,L)
=\frac{
 \left|\left\{u\in\{0,\ldots,R_L-1\}:
 h_T(2+3L+u)=c\right\}\right|
}{R_L}.
\tag{9}
$$
Equation (9) makes the 64-emerald row heavier whenever the cap combines several
underlying values. It also prevents a min--max range from implying prices that
cannot occur.

Combining (6) and (9), the generator probability for one enchanted-book offer
is

$$p_{\mathrm{function}}(e,L,c)
=\frac{1}{39m_e}
\times \frac{N_{e,L,c}}{5+10L},
\tag{10}
$$
where $N_{e,L,c}$ is the number of pre-cap random values producing $c$.
Multiplication by the source-trade selection probability completes (3).

### A fully unlocked librarian

Let $j=1,\ldots,4$ index the four book opportunities, and let $A$ be any
query-defined set of acceptable enchantment, level, and price outcomes. The
qualifying probability for opportunity $j$ is

$$q_j(A)
=s_j
\times \sum_{(e,L,c)\in A}
\frac{1}{39m_e}
\times \frac{N_{e,L,c}}{5+10L},
\tag{11}
$$
with $s=(1/2,1/2,1/2,1/3)$. Treating the four group selections as independent,
the probability that a fully unlocked librarian offers at least one
qualifying book is

$$\Pr(\text{at least one qualifying book})
=1-\prod_{j=1}^{4}\{1-q_j(A)\}.
\tag{12}
$$
This complement form avoids double counting librarians whose inventories
contain two or more qualifying books.

## Enchanted equipment

Armorer, fisherman, fletcher, toolsmith, and weaponsmith tables contain 16
`enchant_with_levels` trades. Microsoft describes the function as applying the
enchanting-table algorithm with an input level range
[@microsoft2025enchanting]. In the pinned trades, treasure enchantments are
disabled and the source level is uniform from 5 through 19. bedrocktrader
enumerates the complete distribution analytically rather than simulating it.

### Source level, price, and base power

Let

$$S\sim\operatorname{DiscreteUniform}\{5,\ldots,19\}.
\tag{13}
$$
If the authored base emerald cost is $b$, the exact displayed equipment price
is

$$C=b+S.
\tag{14}
$$
This dependence is important: price and enchantment set share the same source
level. Aggregating enchantment sets first and attaching an independent price
range would destroy that dependence and give incorrect probabilities under an
emerald cutoff.

Let $A_i$ be the item's enchantability and
$r_i=\lfloor A_i/4\rfloor$. Two independent rolls satisfy

$$X_1,X_2
\overset{\mathrm{iid}}{\sim}
\operatorname{DiscreteUniform}\{0,\ldots,r_i\}.
\tag{15}
$$
The base enchanting power is

$$B=S+1+X_1+X_2.
\tag{16}
$$
Next, let $M$ follow the symmetric triangular distribution on
$[0.85,1.15]$ with mode 1. Its cumulative distribution function is

$$F_M(m)=
\begin{cases}
0, & m\leq0.85,\\
\dfrac{(m-0.85)^2}{2(0.15)^2}, & 0.85<m\leq1,\\
1-\dfrac{(1.15-m)^2}{2(0.15)^2}, & 1<m<1.15,\\
1, & m\geq1.15.
\end{cases}
\tag{17}
$$
The integer modified power is

$$P=\left\lfloor BM+\frac{1}{2}\right\rfloor.
\tag{18}
$$
For fixed $B$, its probability mass follows directly from the CDF:

$$\Pr(P=p\mid B)
=F_M\left(\frac{p+1/2}{B}\right)
-F_M\left(\frac{p-1/2}{B}\right).
\tag{19}
$$
The updater averages (19) over the 15 source levels and the
$(r_i+1)^2$ equally likely enchantability-roll pairs. This analytic
integration preserves small probabilities that a finite simulation could miss.

### Eligible levels and weighted selection

Each enchantment $e$ has a weight $w_e$, item-compatibility rule, and power
interval $[a_{e,\ell},b_{e,\ell}]$ for level $\ell$. At power $P$, the
eligible level is the highest $\ell$ satisfying

$$a_{e,\ell}\leq P\leq b_{e,\ell}.
\tag{20}
$$
If no level satisfies (20), the enchantment is absent from the candidate set.
Treasure enchantments are removed because every pinned equipment function sets
`treasure = false`. The first enchantment is drawn from the eligible set
$\mathcal{E}(P)$ with

$$\Pr(e\text{ selected first}\mid P)
=\frac{w_e}{\sum_{a\in\mathcal{E}(P)}w_a}.
\tag{21}
$$
After an enchantment is selected, it and all conflicting enchantments are
removed. The implemented conflict families are the four protection
enchantments, the three damage enchantments, Fortune versus Silk Touch, and
Multishot versus Piercing. Accordingly, the package never returns complete
sets containing both members of those incompatible families.

### Additional enchantments

At a recursive step with current power $P_r$, the probability of attempting
another enchantment is

$$\rho(P_r)
=\min\left\{\frac{P_r+1}{50},1\right\}.
\tag{22}
$$
If the process continues, the power becomes

$$P_{r+1}=\left\lfloor\frac{P_r}{2}\right\rfloor,
\tag{23}
$$
and one of the remaining compatible candidates is selected by its relative
weight. If none remain, the current set is final. Otherwise, termination at
that step has probability $1-\rho(P_r)$.

For a particular ordered selection path
$\pi=(e_1,\ldots,e_m)$, the conditional probability is the product of its
relative selection weights, continuation probabilities through step $m-1$,
and the final stopping probability:

$$\Pr(\pi\mid P)
=
\left[
\prod_{r=1}^{m}
\frac{w_{e_r}}{\sum_{a\in\mathcal{E}_r}w_a}
\right]
\left[
\prod_{r=1}^{m-1}\rho(P_r)
\right]
\tau_m,
\tag{24}
$$
where $\mathcal{E}_r$ is the compatible candidate set at step $r$, and
$\tau_m=1-\rho(P_m)$ unless no candidates remain, in which case
$\tau_m=1$. Different paths can finish with the same unordered complete set.
bedrocktrader sorts its namespaced `identifier=level` pairs and sums all paths
sharing that representation:

$$\Pr(H=h\mid S)
=\sum_{B,P,\pi:\,\operatorname{set}(\pi)=h}
\Pr(B\mid S)\Pr(P\mid B)\Pr(\pi\mid P).
\tag{25}
$$
Together, (13), (14), and (25) define the exact equipment offer:

$$\Pr(H=h,C=c)
=
\sum_{s=5}^{19}
\frac{1}{15}
\times\mathbf{1}\{c=b+s\}
\times\Pr(H=h\mid S=s).
\tag{26}
$$
The source-trade and explicit-choice factors in (3) are then applied.
Option-view rows sum (26) over $c$, whereas offer-view rows retain each
price-specific term.

## Query probabilities

### Enchanted books

`enchanted_book_probability()` defines $A$ in (11) using one requested
`identifier=level` pair and an inclusive original-price cutoff $C_{\max}$.
With `include_higher_level = FALSE`, an outcome qualifies when
$$e=e^\star,\qquad L=L^\star,\qquad C\leq C_{\max}.
$$

When `include_higher_level = TRUE`, the level requirement becomes
$L\geq L^\star$. The function then applies (12). The cutoff concerns the
modeled original offer and does not simulate curing.

### Enchanted equipment

For equipment, write the requested set as
$$Q=\{(e_1,\ell_1),\ldots,(e_r,\ell_r)\}
$$

and an observed complete set as $H$. Under `match = 'exact'`, the identifiers
in $H$ must equal those in $Q$. Under `match = 'contains'`, $Q$'s
identifiers need only be a subset of $H$. The level comparison is equality by
default and becomes $L_H(e)\geq\ell_e$ when
`include_higher_level = TRUE`.

Let $\mathcal{Q}$ be the offer rows that satisfy the identifier rule, level
rule, item rule, profession rule, and $C\leq C_{\max}$. The returned
probability is

$$\Pr(\mathcal{Q})
=\sum_{o\in\mathcal{Q}}p_{\mathrm{offer}}(o\mid c).
\tag{27}
$$
The events are disjoint exact offers, so direct summation is appropriate.
Diamond axes require a profession because both toolsmiths and weaponsmiths
sell them. Other supported equipment items imply a profession. Recognized but
item-inapplicable enchantments, incompatible requests, and unavailable
enchantments define an empty $\mathcal{Q}$ and return zero.

## Worked examples

### Mending for at most 26 emeralds

Mending is a level-I treasure enchantment. From (7), its 15 equally likely
ordinary underlying prices run from 5 through 19. After (8), the displayed
support is $10,12,\ldots,38$. Nine of those 15 values are at most 26, so the
conditional price probability is $9/15=3/5$. Within a selected book trade,

$$\Pr(\text{Mending I and }C\leq26)
=\frac{1}{39}\frac{3}{5}
=\frac{1}{65}.
$$
The four opportunity probabilities are therefore
$$q_1=q_2=q_3=\frac{1}{2}\frac{1}{65}=\frac{1}{130},
\qquad
q_4=\frac{1}{3}\frac{1}{65}=\frac{1}{195}.
$$
Equation (12) gives

$$1-\left(1-\frac{1}{130}\right)^3
  \left(1-\frac{1}{195}\right)
=0.027910633\ldots,
$$
or approximately $2.79\%$. The corresponding R call is:

```{r}
# Calculate Mending probability.
mending_probability <- enchanted_book_probability(
  enchantment  = 'minecraft:mending=1',
  max_emeralds = 26
)
```

The often-used 26-emerald threshold is a screening rule for a favorable
original librarian price. The package does not claim that every such book will
become a one-emerald trade, because curing and subsequent price adjustments are
outside this model.

### Exact Efficiency II diamond axes

The next query requires the complete enchantment set to be exactly Efficiency
II, permits every original price, and does not count higher Efficiency levels:

```{r}
# Calculate exact Efficiency II diamond axe probabilities.
toolsmith_probability <- enchanted_item_probability(
  item         = 'axe',
  enchantments = 'efficiency=2',
  profession   = 'toolsmith'
)

weaponsmith_probability <- enchanted_item_probability(
  item         = 'axe',
  enchantments = 'efficiency=2',
  profession   = 'weaponsmith'
)
```

Analytic enumeration of (13)--(26) yields
$0.060656412$, or approximately $6.07\%$, for a fully unlocked toolsmith
and $0.121312824$, or approximately $12.13\%$, for a fully unlocked
weaponsmith. The conditional enchanting distribution is the same diamond-axe
distribution in both tables. Their factor-of-two difference comes from source
selection: the toolsmith axe trade is selected with probability $1/2$, while
the weaponsmith axe trade is guaranteed.

Changing `match = 'contains'` would also admit axes that carry Efficiency II
alongside compatible enchantments. Setting `include_higher_level = TRUE` would
admit Efficiency III and above as well.

### Candle multiplicity and probability conservation

The candle group demonstrates the entire hierarchy without an enchanting
function. At trade resolution, three red source entries collapse into one row
with probability $3/4$; yellow has probability $1/4$. Each color is already
a concrete specification and fixed price, so its option and offer probability
remain unchanged:

$$p_{\mathrm{trade}}(\text{red})
=p_{\mathrm{option}}(\text{red})
=p_{\mathrm{offer}}(\text{red})
=\frac{3}{4}.
$$
For enchanted equipment or books, one option contains several prices. Equation
(4) becomes an observable conservation check:

$$\sum_{\substack{o:\,
\texttt{option\_id}(o)=a}}
p_{\mathrm{offer}}(o)
=p_{\mathrm{option}}(a).
$$
Summing those options according to (5) recovers the corresponding trade-view
probability. The package's lightweight tests exercise these identities for all
armorer views.

## Model boundaries

The model deliberately separates source-resolved quantities from documented
reconstructions and unresolved engine behavior. `probability_status = 'exact'`
means the result follows directly from the pinned table's selection, choices,
and fixed or uniformly enumerated functions. `documented_model` means the
probability is exact under the mathematical model in this note. It does not
elevate Minecraft Wiki mechanics to an official Mojang guarantee.
`random_dye` remains `partial`; bedrocktrader does not invent a distribution
for leather colors when the pinned sources do not establish one.

All emerald cutoffs refer to the original modeled price. Demand, depleted
stock, player reputation, curing discounts, and Hero of the Village effects can
alter the amount shown in game. Those mechanisms operate after the trade
outcome modeled here and are outside the package's current probability
functions. The exposed `price_multiplier` remains useful for identifying
trade-table behavior, but it is not used to forecast a later price.

The librarian model assumes the documented uniform enchantment-then-level
selection among 39 attainable enchantments. The equipment model excludes
treasure enchantments because the pinned `enchant_with_levels` entries do so.
It also depends on the documented enchantability values, eligible power
intervals, weights, conflicts, continuation rule, and triangular modifier in
the pinned reference revision. A future Bedrock release can change identifiers,
trade structures, or engine rules; this package therefore supports only
`1.26.30.5` until the model can be reviewed against another complete release.

The updater fails closed when a pinned JSON shape, function form, filter, or
checksum differs from what the parser understands. No Mojang JSON is
redistributed verbatim. The normalized values remain subject to Mojang's
license notice and the Minecraft End User License Agreement
[@mojang2026license]. bedrocktrader's R code is distributed under GPL-3-or-later.

## Conclusion

Villager trading combines several modest random mechanisms whose interaction is
easy to misread from a flattened table. Conditioning on context determines the
applicable source pool; selection without replacement produces source-trade
probabilities; multiplicity combines identical authored paths; item choices
and functions generate conditional branches; and summation connects exact
offers to item options and trade summaries. Librarian books add a
uniform-enchantment, uniform-level, and capped-price model, while equipment
trades couple a source level to both emerald price and a recursive enchanting
algorithm.

The `trade_id -> option_id -> offer_id` hierarchy keeps those resolutions
visible. It supports concise trade inspection, exact probability queries, and
conservation checks without simulation. Most importantly, each public
probability answers the question posed by its row: a summarized trade, a
specified item over possible prices, or one exact offer.

## References
